---
title: "Comparison fire epoch (Congreso Pastos 2023)"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, message=FALSE)
```

# Introduction

```{r pkg}
library(tidyverse)
library(here)
library(janitor)
library(mgcv)
library(gratia)
library(itsadug)
library(kableExtra)
library(tidymv)
library(broom)
library(performance)
library(patchwork)
library(gtsummary)
library(plotrix)
```

- Read data 

```{r read-data, echo=TRUE}
veg_raw <- readxl::read_excel(here::here("data/datos_sep2023_todos_juntos.xlsx")) |> 
  janitor::clean_names() 

autumn_fire <- lubridate::ymd("2018-12-18")
spring_fire <- lubridate::ymd("2019-05-07")

veg <- veg_raw |> 
  filter(zona != "Quemado sin pastoreo") |> 
  rename(fecha = fecha_muestreos) |> 
  mutate(treat = as.factor(case_when(
    zona == "Quemado con pastoreo" ~ "Ot", 
    TRUE ~ "Pr"
  ))) |> 
  mutate(meses = ifelse(treat == "Ot",
                        as.numeric(as.character(lubridate::interval(autumn_fire, lubridate::ymd(fecha)) %/% months(1))),
                        as.numeric(as.character(lubridate::interval(spring_fire, lubridate::ymd(fecha)) %/% months(1))))) |> 
  mutate(quadrat = as.factor(quadrat)) |> 
  mutate(meses = case_when(
    meses == 0 ~ -1, 
    TRUE ~ meses
  ))

veg_total <- veg |> filter(tipo == "total")
veg_mt <- veg |> filter(tipo == "Macrochloa tenacissima") 
veg_gs <- veg |> filter(tipo == "Genista scorpius") 

# Subset of data (pre/post) 
ad_total <- veg_total |> filter(meses %in% c(-1, 22, 24)) 
ad_mt <-  veg_mt |> filter(meses %in% c(-1, 22, 24)) 
ad_gs <- veg_gs |> filter(meses %in% c(-1, 22, 24))
```


```{r auxR}
# Modify the itsadug::gamtab, to export as data.frame
gamtab2 <- function (model, caption = " ", label = "tab.gam", pnames = NA, 
    snames = NA, ptab = NA, stab = NA, ...) 
{
    if (!requireNamespace("xtable", quietly = TRUE)) {
        stop("Package 'xtable' needed for this function to work. Please install it.", 
            call. = FALSE)
    }
    sum.gam <- model
    if (!inherits(model, "summary.gam")) {
        sum.gam <- summary(model)
    }
    if (is.na(ptab[1])) {
        ptab = as.data.frame(sum.gam$p.table, stringsAsFactors = FALSE)
    }
    if (is.na(stab[1])) {
        stab = as.data.frame(sum.gam$s.table, stringsAsFactors = FALSE)
    }
    if (!is.na(pnames[1])) {
        rownames(ptab) = pnames
    }
    if (!is.na(snames[1])) {
        rownames(stab) = snames
    }
    colnames(ptab)[4] = "p-value"
    colnames(ptab)[3] = "t-value"
    ptab.cnames = colnames(ptab)
    stab.cnames = colnames(stab)
    stab.cnames[3] = "F-value"
    colnames(ptab) = c("A", "B", "C", "D")
    if (ncol(stab) != 0) {
        colnames(stab) = colnames(ptab)
    }
    tab = rbind(ptab, stab)
    colnames(tab) = ptab.cnames
    tab = round(tab, 4)
    m = data.frame(matrix(0, nrow(tab), ncol(tab)), stringsAsFactors = FALSE)
    for (i in 1:nrow(tab)) {
        for (j in 1:4) {
            if ((j == 4) & (tab[i, j] < 1e-04)) {
                m[i, j] = "< 0.0001"
            }
            else {
                m[i, j] = sprintf("%3.4f", tab[i, j])
            }
        }
    }
    colnames(m) = colnames(tab)
    rownames(m) = rownames(tab)
    tab = m
    tab2 = rbind(c(ptab.cnames), tab[1:nrow(ptab), ])
    if (nrow(stab) > 0) {
        tab2 = rbind(tab2, c(stab.cnames), tab[(nrow(ptab) + 
            1):nrow(tab), ])
    }
    if (nrow(stab)) {
        rownames(tab2)[(nrow(ptab) + 2)] = "B. smooth terms"
    }
    rownames(tab2)[1] = "A. parametric coefficients"
    for (i in 1:nrow(tab2)) {
        if (tab2[i, 4] == "0") 
            tab2[i, 4] = "< 0.0001"
        if (length(grep("\\.", tab2[i, 2])) == 0) 
            tab2[i, 2] = paste(tab2[i, 2], ".0000", sep = "")
    }

return(as.data.frame(tab2))
}



gamR2 <- function(gam) {
  # see https://www.researchgate.net/post/How-can-I-get-the-adjusted-r-squared-value-of-GAM-model
  R2 <- 1 - ((sum(residuals(gam)^2)) /
    (sum((gam$y - mean(gam$y))^2)))
  R2adj <- 1 - ((1 - R2) * (length(gam$y) - 1) /
    (length(gam$y) - length(gam$coefficients) - 1))
  a <- data.frame(R2, R2adj)
  return(a)
}


library(emmeans)
tabla_postHoc <- function(modelo){
  posthoc <- emmeans(modelo, ~ meses * treat)
  contrast(posthoc, "consec", simple = "each", combine = TRUE,
           adjust = "tukey") |> 
  kbl(digits = 4) |> 
    kable_paper("hover", full_width = F)
}


colores <- c("#543005", "#01665e") 

```

# All sps 
## Total Cover 
### Model
```{r, echo=TRUE}
f <- as.formula(cob_total ~ s(meses, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg_total, 
          family = nb)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.rec_tot <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Negative Binomial")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.rec_tot_all <- broom::tidy(m$gam) |> mutate(var = "Total cover", 
                                                   tipo = "All species") 
param.rec_tot_all <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Total cover", 
                                                                     tipo = "All species")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.rec_total <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Recubrimiento total (%)") + 
  xlab("Tiempo (meses tras la quema)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_total, 
             aes(x=meses, y=cob_total, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) 


inset.rec_tot <- ggpubr::ggboxplot(ad_total, x="meses", y="cob_total", fill = "treat", palette = colores, ylab="") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))
```

```{r plot_rectot, dev = c('png','jpeg','tiff')}
plot.rec_total + inset_element(inset.rec_tot,
                         0.4, 0.5, 1,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_total, 
  x=meses,
  y=cob_total, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(cob_total~meses*treat, data = ad_total)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```

## Phytovol
### Model
```{r, echo=TRUE}
f <- as.formula(fitovol ~ s(meses, k=5, bs="cs") + treat)
m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg_total, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.fitovol1 <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Tweedie")
```


### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.fitovol1_all <- broom::tidy(m$gam) |> mutate(var = "Fitovol", tipo = "All species")
param.fitovol1_all <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Fitovol", tipo = "All species")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.fitovol <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat,
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(Fitovolumen~(m^3~ha^{-1}))) + 
  xlab("Tiempo (meses tras la quema)") +  
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_total, 
             aes(x=meses, y=fitovol, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) 

inset.fitovol <- ggpubr::ggboxplot(ad_total, x="meses", y="fitovol", fill = "treat", palette = colores, ylab="") + 
  theme(legend.position = "none", 
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))

```

```{r plot_fitovol, dev = c('png','jpeg','tiff')}
plot.fitovol + inset_element(inset.fitovol, 0.4, 0.4, .95,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_total, 
  x=meses,
  y=fitovol, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(fitovol~meses*treat, data = ad_total)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```


# Macrochloa tenacissima
## Total Cover 
### Model
```{r, echo=TRUE}
f <- as.formula(cob_total ~ s(meses, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg_mt, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.rec_tot <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Tweedie")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.rec_tot_mt <- broom::tidy(m$gam) |> mutate(var = "Total cover", tipo = "Macrochloa tenacissima")
param.rec_tot_mt <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Total cover", tipo = "Macrochloa tenacissima")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.rec_total_mt <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Recubrimiento total (%)") + 
  xlab("Tiempo (meses tras la quema)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_mt, 
             aes(x=meses, y=cob_total, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) +
  ggtitle("Macrochloa tenacissima")



inset.rec_tot_mt <- ggpubr::ggboxplot(ad_mt, x="meses", y="cob_total", fill = "treat", palette = colores, ylab="") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))
```

```{r plot_rectot_mt, dev = c('png','jpeg','tiff')}
plot.rec_total_mt + inset_element(inset.rec_tot_mt,
                         0.4, 0.5, 1,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_mt, 
  x=meses,
  y=cob_total, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(cob_total~meses*treat, data = ad_mt)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```

## Phytovol
### Model
```{r, echo=TRUE}
f <- as.formula(fitovol ~ s(meses, k=5, bs="cs") + treat)
m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg_mt, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.fitovol1 <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Tweedie")
```

### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.fitovol1_mt <- broom::tidy(m$gam) |> mutate(var = "Fitovol", tipo = "Macrochloa tenacissima")
param.fitovol1_mt <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Fitovol", tipo = "Macrochloa tenacissima")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.fitovol_mt <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat,
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(Fitovolumen~(m^3~ha^{-1}))) + 
  xlab("Tiempo (meses tras la quema)") +  
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_mt, 
             aes(x=meses, y=fitovol, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) +
  ggtitle("Macrochloa tenacissima")

inset.fitovol_mt <- ggpubr::ggboxplot(ad_mt, x="meses", y="fitovol", fill = "treat", palette = colores, ylab="") + 
  theme(legend.position = "none", 
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))

```

```{r plot_fitovol_mt, dev = c('png','jpeg','tiff')}
plot.fitovol_mt + inset_element(inset.fitovol_mt, 0.4, 0.4, .95,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_total, 
  x=meses,
  y=fitovol, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(fitovol~meses*treat, data = ad_mt)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```



# Genista scorpius
## Total Cover 
### Model
```{r, echo=TRUE}
f <- as.formula(cob_total ~ s(meses, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg_gs, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.rec_tot <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Tweedie")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.rec_tot_gs <- broom::tidy(m$gam) |> mutate(var = "Total cover", tipo = "Genista scorpius")
param.rec_tot_gs <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Total cover", tipo = "Genista scorpius")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.rec_total_gs <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Recubrimiento total (%)") + 
  xlab("Tiempo (meses tras la quema)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_gs, 
             aes(x=meses, y=cob_total, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) +
  ggtitle("Genista scorpius")



inset.rec_tot_gs <- ggpubr::ggboxplot(ad_gs, x="meses", y="cob_total", fill = "treat", palette = colores, ylab="") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))
```

```{r plot_rectot_gs, dev = c('png','jpeg','tiff')}
plot.rec_total_gs + inset_element(inset.rec_tot_gs,
                         0.4, 0.5, 1,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_gs, 
  x=meses,
  y=cob_total, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(cob_total~meses*treat, data = ad_gs)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```

## Phytovol
### Model
```{r, echo=TRUE}
f <- as.formula(fitovol ~ s(meses, k=5, bs="cs") + treat)
m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg_gs, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.fitovol1 <- 
  performance(m$gam) |> as.data.frame() |> mutate(dist = "Tweedie")
```

### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) |> 
  kbl(col.names = NULL) |> 
  kable_paper(full_width = F) |>
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.fitovol1_gs <- broom::tidy(m$gam) |> mutate(var = "Fitovol", tipo = "Genista scorpius")
param.fitovol1_gs <- broom::tidy(m$gam, parametric = TRUE) |> mutate(var = "Fitovol", tipo = "Genista scorpius")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
plot.fitovol_gs <- plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat,
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(Fitovolumen~(m^3~ha^{-1}))) + 
  xlab("Tiempo (meses tras la quema)") +  
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="black") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg_gs, 
             aes(x=meses, y=fitovol, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter(width = .4)) +
  ggtitle("Genista scorpius")

inset.fitovol_gs <- ggpubr::ggboxplot(ad_gs, x="meses", y="fitovol", fill = "treat", palette = colores, ylab="") + 
  theme(legend.position = "none", 
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        text = element_text(size = 10))

```

```{r plot_fitovol_gs, dev = c('png','jpeg','tiff')}
plot.fitovol_gs + inset_element(inset.fitovol_gs, 0.4, 0.4, .95,.95)
```

### Pre-post (autumn)
```{r}
ggstatsplot::grouped_ggwithinstats(
  ad_total, 
  x=meses,
  y=fitovol, 
  grouping.var = treat,
  pairwise.display = "significant",
  point.path = FALSE,
  type = "parametric"
)
```

```{r}
maov <- aov(fitovol~meses*treat, data = ad_gs)
```

```{r}
options(knitr.kable.NA = '')
broom::tidy(maov) |> 
  kbl(digits = c(0, 0, 2,2,2,4)) |> 
  kable_styling()
```












# Summary 

- Parametric terms 

```{r}
p.terms <- 
  bind_rows(param.rec_tot_all,
            param.fitovol1_all,
            param.rec_tot_mt, 
            param.fitovol1_mt,
            param.rec_tot_gs, 
            param.fitovol1_gs) |> 
  relocate(tipo, var) |> 
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "treatOt", 
      TRUE ~ term)
    )
```


```{r}
p.terms |> 
  mutate(p = 
           case_when(
             p.value < 0.0001 ~ "< 0.0001",
             TRUE ~ as.character(round(p.value,4))
           )) |> 
  dplyr::select(-p.value) |> 
  kbl(col.names = 
        c("Tipo", "Variable", "term", "Estimate", "SE", "F", "p"), 
      align = c("ccrrrr"),
      digits = c(0,0,0,3,3,2,4)) |> 
  kable_paper(full_width = F) 
```


- Smooth terms
```{r}
s.terms <- 
  bind_rows(smooth.rec_tot_all,
            smooth.fitovol1_all,
            smooth.rec_tot_mt, 
            smooth.fitovol1_mt,
            smooth.rec_tot_gs, 
            smooth.fitovol1_gs) |> 
  relocate(tipo, var) 
```

```{r}
s.terms |> 
  mutate(p = 
           case_when(
             p.value < 0.0001 ~ "< 0.0001",
             TRUE ~ as.character(p.value)
           )) |> 
  dplyr::select(-p.value) |> 
  kbl(col.names = 
        c("tipo", "Variable", "term", "edf", "ref.df", "F", "p"), 
      digits = c(0,0,0,3,3,2,4)) |> 
  kable_paper(full_width = F) 
```



# All plots 

```{r}
library(patchwork)
```

```{r, fig.height=10}
plot.rec_total + plot.fitovol + 
  plot.rec_total_mt + plot.fitovol_mt +
  plot.rec_total_gs + plot.fitovol_gs +
  plot_layout(guides = "collect", nrow = 3) & theme(legend.position = "bottom")
```


# Summary table 

```{r}
veg_total |> 
  dplyr::select(cob_total, fitovol, treat, meses) |> 
 tbl_strata(
    strata = treat, 
    .tbl_fun = 
      ~.x |> 
      tbl_summary(
        by = meses,
        missing = "no",
        # type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ± {std.error}",
        digits = everything() ~ 2
      )) |> 
    modify_caption("**All species**")
```

```{r}
veg_mt |> 
  dplyr::select(cob_total, fitovol, treat, meses) |> 
 tbl_strata(
    strata = treat, 
    .tbl_fun = 
      ~.x |> 
      tbl_summary(
        by = meses,
        missing = "no",
        # type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ± {std.error}",
        digits = everything() ~ 2
      )) |> 
  modify_caption("*Macrochloa tenacissima*")
```


```{r}
veg_gs |> 
  dplyr::select(cob_total, fitovol, treat, meses) |> 
 tbl_strata(
    strata = treat, 
    .tbl_fun = 
      ~.x |> 
      tbl_summary(
        by = meses,
        missing = "no",
        # type = c(p) ~ "continuous",
       # label = list(ab ~ "Juvenile abundance"),
        statistic = all_continuous() ~ "{mean} ± {std.error}",
        digits = everything() ~ 2
      )) |> 
  modify_caption("*Genista scorpius*")
```





