---
title: "temporal_analysis"
author: "ajpelu"
date: "2021-10-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

# Introduction

```{r pkg}
library(tidyverse)
library(here)
library(janitor)
library(mgcv)
library(gratia)
library(itsadug)
library(kableExtra)
library(tidymv)
library(broom)
library(performance)
```

- Read data 
- Rename the levels of *zona*: 

  - **Pyric herbivorism** (***HP***, from spanish *Herbivorismo pírico*): old *Quemado con pastoreo*
  - **Prescribed Fires** (***QP***, from spanish *Quemas Prescritas*): old *Quemado sin pastoreo*
  

```{r read-data, echo=TRUE}
veg_raw <- readxl::read_excel(here::here("data/Datos_congreso_AEET.xlsx")) %>% 
  janitor::clean_names() 

autumn_fire <- lubridate::ymd("2018-12-18")
grazing_start <- lubridate::ymd("2019-10-01")

veg <- veg_raw %>% 
  rename(fecha = fecha_muestreos, 
         shan = i_shannon, 
         consumo = tasa_consumo, 
         riq = riqueza) %>% 
  mutate(treat = as.factor(case_when(
    zona == "Quemado con pastoreo" ~ "HP", 
    TRUE ~ "QP"
  ))) %>% 
  mutate(meses = as.numeric(as.character(lubridate::interval(autumn_fire, lubridate::ymd(fecha)) %/% months(1)))) %>% 
  mutate(quadrat = as.factor(quadrat)) %>% 
  mutate(meses = case_when(
    meses == 0 ~ -1, 
    TRUE ~ meses
  ))
 
```

  
- Explore the data 

```{r}
xtabs(~treat+fecha, data = veg)
```
 
# Statistical analysis

- We used Generalizad Additive Mixed Models (GAMM) to evaluate the effects of treatment (Prescribed Fires *vs.* Pyric herbivorism) on the evolution of plant cover, phytovolumen and diversity indices (Shannon, Richness). 

- We modelling the date after fire (in months; *meses*) as smoother (`s(meses)`)

- Date after fire and treatment was include as fixed effects 

- Quadrats were included as random 

- Different distribution family (*e.g.* Poisson, Negative Binomial, etc) were used. 

# Richness
## Model
```{r gamm-riq, echo=TRUE}
f <- as.formula(riq ~ s(meses, by = treat, k=5, bs="cs") + treat)

m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg, 
          family = poisson)
```

```{r, echo=FALSE}
performance::check_distribution(m$gam)
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r auxR}
# Modify the itsadug::gamtab, to export as data.frame
gamtab2 <- function (model, caption = " ", label = "tab.gam", pnames = NA, 
    snames = NA, ptab = NA, stab = NA, ...) 
{
    if (!requireNamespace("xtable", quietly = TRUE)) {
        stop("Package 'xtable' needed for this function to work. Please install it.", 
            call. = FALSE)
    }
    sum.gam <- model
    if (!inherits(model, "summary.gam")) {
        sum.gam <- summary(model)
    }
    if (is.na(ptab[1])) {
        ptab = as.data.frame(sum.gam$p.table, stringsAsFactors = FALSE)
    }
    if (is.na(stab[1])) {
        stab = as.data.frame(sum.gam$s.table, stringsAsFactors = FALSE)
    }
    if (!is.na(pnames[1])) {
        rownames(ptab) = pnames
    }
    if (!is.na(snames[1])) {
        rownames(stab) = snames
    }
    colnames(ptab)[4] = "p-value"
    colnames(ptab)[3] = "t-value"
    ptab.cnames = colnames(ptab)
    stab.cnames = colnames(stab)
    stab.cnames[3] = "F-value"
    colnames(ptab) = c("A", "B", "C", "D")
    if (ncol(stab) != 0) {
        colnames(stab) = colnames(ptab)
    }
    tab = rbind(ptab, stab)
    colnames(tab) = ptab.cnames
    tab = round(tab, 4)
    m = data.frame(matrix(0, nrow(tab), ncol(tab)), stringsAsFactors = FALSE)
    for (i in 1:nrow(tab)) {
        for (j in 1:4) {
            if ((j == 4) & (tab[i, j] < 1e-04)) {
                m[i, j] = "< 0.0001"
            }
            else {
                m[i, j] = sprintf("%3.4f", tab[i, j])
            }
        }
    }
    colnames(m) = colnames(tab)
    rownames(m) = rownames(tab)
    tab = m
    tab2 = rbind(c(ptab.cnames), tab[1:nrow(ptab), ])
    if (nrow(stab) > 0) {
        tab2 = rbind(tab2, c(stab.cnames), tab[(nrow(ptab) + 
            1):nrow(tab), ])
    }
    if (nrow(stab)) {
        rownames(tab2)[(nrow(ptab) + 2)] = "B. smooth terms"
    }
    rownames(tab2)[1] = "A. parametric coefficients"
    for (i in 1:nrow(tab2)) {
        if (tab2[i, 4] == "0") 
            tab2[i, 4] = "< 0.0001"
        if (length(grep("\\.", tab2[i, 2])) == 0) 
            tab2[i, 2] = paste(tab2[i, 2], ".0000", sep = "")
    }

return(as.data.frame(tab2))
}



gamR2 <- function(gam) {
  # see https://www.researchgate.net/post/How-can-I-get-the-adjusted-r-squared-value-of-GAM-model
  R2 <- 1 - ((sum(residuals(gam)^2)) /
    (sum((gam$y - mean(gam$y))^2)))
  R2adj <- 1 - ((1 - R2) * (length(gam$y) - 1) /
    (length(gam$y) - length(gam$coefficients) - 1))
  a <- data.frame(R2, R2adj)
  return(a)
}
```

```{r tab-modriq}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)


# Get parameters 
smooth.riq <- broom::tidy(m$gam) %>% mutate(var = "Richness")
param.riq <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Richness")
```

```{r}
performance.riq <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Poisson")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
 transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Richness") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, aes(x=meses, y=riq, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
  
```



# Shannon 
## Model
```{r, echo=TRUE}
veg$shannon <- abs(veg$shan)

f <- as.formula(shannon ~ s(meses, by = treat, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.shannon <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Tweedie")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.shannon<- broom::tidy(m$gam) %>% mutate(var = "Shannon")
param.shannon <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Shannon")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Shannon") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, 
             aes(x=meses, y=shannon, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
  
```



# Total Cover 
## Model
```{r, echo=TRUE}
f <- as.formula(rec_tot ~ s(meses, by = treat, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg, 
          family = nb)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.rec_tot <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Negative Binomial")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.rec_tot <- broom::tidy(m$gam) %>% mutate(var = "Total cover")
param.rec_tot <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Total cover")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Total Cover (%)") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, 
             aes(x=meses, y=rec_tot, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
  
```


# Vegetation Cover 
## Model
```{r echo=TRUE}
f <- as.formula(rec_veg ~ s(meses, by = treat, k=5, bs="cs") + treat)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg, 
          family = tw) 
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.rec_veg <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Tweedie")
```

### Model validation
```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients
```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.rec_veg <- broom::tidy(m$gam) %>% mutate(var = "Veg. cover")
param.rec_veg <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Veg. cover")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat, 
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Veg Cover (%)") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, 
             aes(x=meses, y=rec_veg, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
```

# Phytovol 
## Model
```{r echo=TRUE}
veg$fitovolT <- sqrt(veg$fitovol)
f <- as.formula(fitovolT ~ s(meses, by = treat, k=5, bs="cs") + treat)
m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.fitovol <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Tweedie")
```


### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.fitovol<- broom::tidy(m$gam) %>% mutate(var = "Fitovol")
param.fitovol <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Fitovol")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat,
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab(expression(sqrt(fitol))) + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, 
             aes(x=meses, y=fitovolT, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
  
```


# Phytovol (I) 
## Model
```{r echo=TRUE}

f <- as.formula(fitovol ~ s(meses, by = treat, k=5, bs="cs") + treat)
m <- gamm(f,
          random = list(quadrat = ~1), 
          data = veg, 
          family = tw)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.fitovol1 <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Tweedie")
```


### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.fitovol1<- broom::tidy(m$gam) %>% mutate(var = "Fitovol")
param.fitovol1 <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Fitovol")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat,
  transform = exp
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Phytovol") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg, 
             aes(x=meses, y=fitovol, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
  
```


# Tasa consumo 
## Model
- Caution, there is an outlier . 

```{r, echo=TRUE}
veg %>% filter(consumo > 8)
```


```{r echo=TRUE}
f <- as.formula(consumo ~ s(meses, by = treat, k=5, bs="cs") + treat)

veg.out <- veg %>% filter(consumo < 8)

m <- gamm(f, 
          random = list(quadrat = ~1), 
          data = veg.out)
```

```{r}
performance::check_distribution(m$gam)
```

```{r}
performance.consumo <- 
  performance(m$gam) %>% as.data.frame() %>% mutate(dist = "Gaussian")
```

### Model validation

```{r}
appraise(m$gam, point_col = "steelblue", point_alpha = 0.4)
```

### Coefficients

```{r}
gamtab2(m$gam) %>% 
  kbl(col.names = NULL) %>% 
  kable_paper(full_width = F) %>%
  row_spec(c(1,4), bold = T)

# Get parameters 
smooth.consumo <- broom::tidy(m$gam) %>% mutate(var = "Tasa Consumo")
param.consumo <- broom::tidy(m$gam, parametric = TRUE) %>% mutate(var = "Tasa Consumo")
```

### Visualizing effects
```{r}
gratia::draw(m$gam)
```

```{r}
colores <- c("#488f31", "#f18255")

plot_smooths(
  model = m$gam,
  series = meses, 
  comparison = treat
) +
  scale_colour_manual(values = colores) +
  scale_fill_manual(values = colores) +
  ylab("Consumo") + 
  xlab("Time (months)") + 
  theme_bw() + 
  theme(legend.position = "top",
        panel.grid = element_blank()) +
  geom_vline(xintercept = 0.5, color="#800909") + 
  geom_vline(xintercept = 
               (lubridate::interval(autumn_fire,grazing_start) %/% months(1)),
             linetype="dotted"
             ) + 
  geom_point(data = veg.out, 
             aes(x=meses, y=consumo, colour=treat), 
             alpha=.6, size=.8,
             position = position_jitter())
```


# General overview 

```{r}
p.terms <- 
  bind_rows(param.riq, 
            param.shannon, 
            param.rec_tot,
            param.rec_veg,
            param.fitovol, 
            param.fitovol1,
            param.consumo) %>% 
  relocate(var) %>%
  mutate(
    term = case_when(
      term == "(Intercept)" ~ "treatHP", 
      TRUE ~ term)
    )
```

- Parametric terms 

```{r}
p.terms %>% 
  mutate(p = 
           case_when(
             p.value < 0.0001 ~ "< 0.0001",
             TRUE ~ as.character(round(p.value,4))
           )) %>% 
  dplyr::select(-p.value) %>% 
  kbl(col.names = 
        c("Variable", "term", "Estimate", "SE", "F", "p"), 
      align = c("ccrrrr"),
      digits = c(0,0,3,3,2,4)) %>% 
  kable_paper(full_width = F) 
```


- Smooth terms
```{r}
s.terms <- 
  bind_rows(smooth.riq,
            smooth.shannon, 
            smooth.rec_tot,
            smooth.rec_veg,
            smooth.fitovol, 
            smooth.fitovol1,
            smooth.consumo) %>% 
  relocate(var) 
```

```{r}
s.terms %>% 
  mutate(p = 
           case_when(
             p.value < 0.0001 ~ "< 0.0001",
             TRUE ~ as.character(p.value)
           )) %>% 
  dplyr::select(-p.value) %>% 
  kbl(col.names = 
        c("Variable", "term", "edf", "ref.df", "F", "p"), 
      digits = c(0,0,3,3,2,4)) %>% 
  kable_paper(full_width = F) 
```


# Some resources 

- [Plots for Generalized Additive Models](https://doi.org/10.1080/03610926.2011.628772)
- [Family distributions for GAMs](https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/family.mgcv.html)
- [Tidy Model Visualisation for Generalised Additive Models](https://stefanocoretta.github.io/tidymv/index.html)
- [Course about GAMs](https://noamross.github.io/gams-in-r-course)
- Hierarchical generalized additive models in ecology: an introduction with mgcv [paper](https://peerj.com/articles/6876/)
- [GAMs](https://m-clark.github.io/generalized-additive-models) 




